% Created 2020-12-25 ven. 18:44
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\usepackage{tabularx}
\author{Vijay Gopal Chilkuri}
\date{\today}
\title{QMC Part I}
\hypersetup{
 pdfauthor={Vijay Gopal Chilkuri},
 pdftitle={QMC Part I},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.1 (Org mode 9.5)}, 
 pdflang={English}}
\begin{document}

\maketitle
\tableofcontents


\section{Part I}
\label{sec:org010db07}

\section{Simple power method}
\label{sec:org1c5e2e4}

\subsection{2x2 Matrix}
\label{sec:org5199702}

Consider a Hamiltonian whose matrix form is shown in Eq:\ref{mat2x2}. This Hamiltonian
has two eigenvectors and two eigenvalues. Let the two eigenvectors be
\(\mathbf{u_0}\) and \(\mathbf{u_1}\). Then consider \(\mathbf{\nu}\) any arbitrary trial vector which is
not an eigenvector of Eq:\ref{mat2x2} but belongs to the 2x2 space. Any such trial vector
can always be written in the form shown in Eq:\ref{vecbasis}.

\begin{equation}
\begin{bmatrix}
 v_0 & -t \\
 -t & v_1 \\
\end{bmatrix}
\label{mat2x2}
\end{equation}

\begin{equation}
\label{vecbasis}
\mathbf{\nu} = c_0 \mathbf{u_0} + c_1 \mathbf{u_1}
\end{equation}

The key idea is the realization that the ground state of the Hamiltonian \ref{mat2x2}
given by \(\mathbf{u_0}\) can be extracted from \(\mathbf{\nu}\) by the repeated
application of a filter \(G(H)\). This filter systematically purifies
\(\mathbf{\nu}\) to obtain the ground state \(\mathbf{u_0}\) provided \(c_0 > 0\),
i.e. the trial vector \(\mathbf{\nu}\) has a non-zero projection on the
ground state.

The form of the filter is inspired from the power method where a successive
application of the Hamiltonian followed by the substraction of the residual
leads to a convergent series of vectors. The limiting value of this convergent
series is one of the extremal eigenvectors of the Hamiltonian. Following this, our filter \(G(H)\) can be written as \ref{filter}

\begin{equation}
\label{filter}
\hat{G}(H) = \left ( \mathbf{1} - \tau (\hat{H} - E_T\mathbf{1}) \right)
\end{equation}

The convergent series of vectors is then \(\left\{ \nu^{(0)},  \nu^{(1)},\
\nu^{(2)},\dots,\nu^{(n)}\right\}\) where \(\nu^{(k)}\) is given by \ref{applyg}

\begin{equation}
\label{applyg}
\nu^{(k+1)} = \hat{G}(H)\nu^{(k)}
\end{equation}

Using Eq:\ref{vecbasis}, Eq:\ref{applyg} can be written as \ref{applyg2}

\begin{equation}
\label{applyg2}
\nu^{(k+1)} = c_0 (1-\tau(E_0-E_T))^{(k)}\mathbf{u_0} + c_1 (1-\tau(E_1-E_T))^{(k)}\mathbf{u_1}
\end{equation}

From Eq:\ref{applyg2}, we can see that a repeated application of the filter Eq:\ref{filter}
with a trial guess energy \(E_T\) will result in the series converging
geometrically to either \(E_0\) or \(E_1\) depending on the choice of \(E_T\).

Here we shall show an example using the Hamiltonian given in Eq:\ref{mat2x2}. In the
case of a 2x2 Hamiltonian, and any general trial vector \([c_0,c_1]\), the
recursion relations for the calculation of \(c^{(k)}_0\) and \(c^{(k)}_1\) are
straight forward and given by Eq:\ref{receqn1},\ref{receqn2}.

\begin{equation}
\label{receqn1}
c^{(k+1)}_0 =  \left(\mathbf{1}-\tau\left(\nu_0 - E_T\right)\right)c^{(k)}_0 + \tau t c^{(k)}_1
\end{equation}

\begin{equation}
\label{receqn2}
c^{(k)}_1 = \tau t c^{(k)}_0 + \left(\mathbf{1}-\tau\left(\nu_1 - E_T\right)\right)c^{(k)}_1
\end{equation}

These are the working equations. As an example, we begin with the initial set of
values given as follows:

\begin{verbatim}
\nu_0=1
\nu_1=2
t=1
\tau=0.1
E_T=3
c_0=0.7071067811865475
c_1=0.7071067811865475
\end{verbatim}

The iteration can begin with these as starting values. The output is given in
the table below.

\begin{verbatim}
10	0.8325345415688012	0.5539731375209688	0.384482692919926
20	0.8481318105729733	0.5297852696085559	0.3820169520382197
30	0.8502938016675892	0.5263083229872757	0.38196704124139036
40	0.8506000682828394	0.5258132024181487	0.38196603207543756
50	0.8506435939378749	0.5257427851073715	0.3819660116711711
60	0.8506497825205341	0.5257327719457557	0.3819660112586187
70	0.8506506624858821	0.5257313481354616	0.3819660112502773
80	0.8506507876108551	0.5257311456790742	0.38196601125010865
\end{verbatim}
\end{document}
